Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-02-08 text/src/article.tex
Two revisions to second section of background

Scaffolding can be described as the process of moving from a series of
disconnected contigs to complete continuous sequence of all the contigs joined
together. The scaffolding process may not be completed but may still resolve
and join some contigs together or add additional regions of sequence to fill
gaps. In the case of a microbial genome, the process of scaffolding may
include the following steps:

@@

Sequence reads may be produced from either DNA strand during the sequencing
process. Therefore the resulting contigs produced from these sequences may are
likely to be representative from either of the genome strands. Contig
orientation involves correcting the sequence data so that all contigs
represent sequence data from the same DNA strand. In the case of microbes this
orientation is often the 5`-3` strand from the origin of DNA replication.

@@

The source location of each contig from the genome is unknown. Contig ordering
represents determining the order of contigs which most accurately represent
the actual genome sequence. This process can also determine if any contigs
represent extra-genomic DNA such as plasmids and therefore should be separated
from the genomic contigs. Contigs are usually ordered in reference to the
origin of replication where the contig enclosing this region is the first
contigs and subsequent contigs are labelled in the direction of DNA
replication.

@@

Once the order of contigs in the genome sequence is known it is useful to
determine their approximate distance from each other. This allows the size of
the gaps between contigs to be determined. Filling in these regions in with
unknown nucleotide characters allows a draft continuous representation of the
genome to determined.

@@

The final stage of producing a genome sequence is to close and fill gaps in
the sequence. This may require returning to the laboratory to perform
additional sequencing to close gaps or \emph{in silico} methods may be able to
close gaps using existing data. Once all gaps in a scaffold have been closed
the genome may be considered finished. Though additional sequencing may be
performed to `proof' any low coverage regions for erroneous nucleotide
insertions, deletions or substitutions.

>>>>

Scaffolding is the process of joining a series disconnected contigs into
complete continuous sequence. Due to genomic complexity, missing data and
diminishing returns scaffolding may not ultimately produce a final completed
sequence. The process of scaffolding may still however joining a subset of
contigs or additional regions of sequence to fill gaps.

@@

The sequencing process produces reads from either strand of the DNA helix. The
resulting contigs constructed from these reads therefore point in either of two
DNA strand directions. Contig orientation involves reverse complementing
sequences so that all contigs point in the same direction and represent the
same DNA strand. In the case of microbes this orientation is often the 5`-3`
strand leading away from the origin of replication.

@@

Contig ordering determines the placement of contigs to best represent the
actual nucleotides as appearing in the genome sequence. This may also determine
if any contigs represent extra-genomic DNA, such as plasmids, and therefore
should be treated separately from genomic contigs. The contig order is commonly
referenced to the origin of replication where the contig containing the
replication start point is first. All subsequent contigs then follow in the
direction of DNA replication.

@@

Given the order and orientation of contigs determining the distances between
contigs may be useful for estimating genome size. The size of each gaps between
contigs represents the length of an unknown region in the genome. Filling in
these regions in with unknown nucleotide characters (`N') allows a draft
continuous representation of the genome to produced. This draft scaffold
sequence represents both the contigs and unknown regions of sequence between
them.

@@

During the scaffolding process closing and filling sequence gaps improves the
accuracy of the genome scaffold. This may require returning to the laboratory
to perform additional sequencing or using computational methods. Once all
contigs have been joined and gaps in a scaffold closed the genome may be
considered as finished. In the case of organisms or particular interest
additional sequencing may be performed to further `proof' any low coverage
regions or detect erroneous nucleotides resulting from insertions, deletions or
substitutions.
